\begin{titlepage}
	
\begin{center}
	
\vspace*{1.0cm}

{\LARGE
\bfseries Uncertainty Quantification for an \\
\vspace*{0.5cm}
Eckstein-Keane-Wolpin model
}
\\
\vspace*{0.5cm}
{\LARGE
\bfseries with correlated input parameters
}

{\large
\vspace*{4.0cm}
Master Thesis Presented to the\\
\vspace*{0.25cm}
Department of Economics at the\\
\vspace*{0.25cm}
Rheinische Friedrichs-Wilhelm-Universität Bonn\\

\vspace*{2.0cm}
in Partial Fulfillment of the Requirements of the Degree of\\
\vspace*{0.25cm}
Master of Science (M.Sc.)\\

\vspace*{4.0cm}
Supervisor: Prof. Dr. Philipp Eisenhauer\\

\vspace*{2.0cm}
Submitted in February 2020 by:\\
Tobias Stenzel\\
Matriculation Number: 2971049
}

\end{center}

\end{titlepage}

\newpage
\section*{Acknowledgements}

The thesis has greatly benefited from the support of several people.\\

\noindent
I want to thank Prof. Philipp Eisenhauer for the supervision of the thesis. He initialised our joint research efforts in uncertainty quantification and he also organises our platforms for discussions and mutual learning. These are our "UQ reading group" and "OpenSourceEconomics". I am especially grateful for his openness, excellent guidance and frequent feedback that he provided.\\

\noindent
Special thanks to Jano\`{s} Gabler for countless valuable programming lessons, his feedbacks and for challenging my views. I want to also thank him, Tobias Raabe and Prof. Philipp Eisenhauer for building and maintaining the research software \textit{respy} and \textit{estimagic} on which this thesis heavily relies.\\

\noindent
I also want to thank JProf. Lena Janys and Tim Mensinger for their lively participation and contributions in our reading group.\\

\noindent
Lastly, I want to thank my parents, Prof. Roswitha and Norbert Stenzel, for supporting me with all their strength throughout my entire life.


\newpage